---
title: "Trade Analysis Report"
subtitle: "La-Z-Boyz of Summer Fantasy Baseball"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    df_print: kable
params:
  my_team: "My Team"
  give: !r c("Player A")
  receive: !r c("Player B")
date: "`r format(Sys.time(), '%B %d, %Y at %I:%M %p')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-dependencies}
# Source analysis modules
source(here::here("R/utils/scoring.R"))
source(here::here("R/utils/player_linker.R"))
source(here::here("R/utils/keeper_value.R"))
source(here::here("R/utils/salary_rules.R"))
source(here::here("R/utils/serialization.R"))
source(here::here("R/analysis/player_valuation.R"))
source(here::here("R/analysis/trade_analyzer.R"))
```

```{r load-data}
# Load cached data
rosters <- load_rds_safe(here::here("data/cache/rosters.rds"))
standings <- load_rds_safe(here::here("data/cache/standings.rds"))

# Load valuations
valuations_path <- here::here("data/cache/valuations.rds")
valuations <- if (file.exists(valuations_path)) load_rds_safe(valuations_path) else NULL

# Load draft data for prospect analysis
draft_files <- list.files(here::here("data/cache"), pattern = "^draft_\\d+\\.rds$",
                          full.names = TRUE)
draft_data <- if (length(draft_files) > 0) {
  do.call(rbind, lapply(draft_files, load_rds_safe))
} else {
  NULL
}
```

```{r run-analysis}
# Run trade analysis
trade_result <- NULL
analysis_error <- NULL

if (!is.null(valuations) && !is.null(rosters) && !is.null(standings)) {
  tryCatch({
    trade_result <- analyze_trade(
      give = params$give,
      receive = params$receive,
      my_team = params$my_team,
      valuations = valuations,
      rosters = rosters,
      standings = standings,
      draft_data = draft_data
    )
  }, error = function(e) {
    analysis_error <<- conditionMessage(e)
  })
}
```

---

## Report Metadata

| Field | Value |
|-------|-------|
| **Analysis Type** | Trade Evaluation |
| **Team** | `r params$my_team` |
| **Giving** | `r paste(params$give, collapse = ", ")` |
| **Receiving** | `r paste(params$receive, collapse = ", ")` |
| **Generated** | `r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')` |

---

```{r error-check, results='asis'}
if (!is.null(analysis_error)) {
  cat(sprintf("## ❌ Analysis Error\n\n**%s**\n\n", analysis_error))
  cat("Please verify that all player names match the roster data exactly.\n")
  knitr::knit_exit()
}
if (is.null(trade_result)) {
  cat("## ❌ Missing Data\n\n")
  cat("Valuations, roster, or standings data not available. ")
  cat("Run the ingest and valuation pipelines first.\n")
  knitr::knit_exit()
}
```

## Recommendation

```{r recommendation, results='asis'}
rec_icon <- switch(trade_result$recommendation,
  "accept" = "✅",
  "reject" = "❌",
  "counter" = "🔄"
)
rec_label <- switch(trade_result$recommendation,
  "accept" = "ACCEPT — This trade improves your team",
  "reject" = "REJECT — This trade is unfavorable",
  "counter" = "COUNTER — Close but could be better"
)
cat(sprintf("### %s %s\n\n", rec_icon, rec_label))

if (!is.null(trade_result$justification)) {
  cat(trade_result$justification)
}
```

---

## Value Analysis

```{r value-analysis}
value_direction <- if (trade_result$value_diff > 0) {
  "in your favor"
} else if (trade_result$value_diff < 0) {
  "against you"
} else {
  "even"
}

value_df <- data.frame(
  Metric = c("Net Surplus Value", "Direction"),
  Value = c(sprintf("%+.1f", trade_result$value_diff), value_direction),
  stringsAsFactors = FALSE
)
knitr::kable(value_df, caption = "Surplus Value Summary")
```

```{r lopsided-flag, results='asis'}
if (isTRUE(trade_result$is_lopsided)) {
  cat("\n⚠️ **LOPSIDED TRADE WARNING:** The net value difference exceeds 20% of ")
  cat("total trade value. This trade is significantly unbalanced.\n\n")
}
```

---

## Points Impact

```{r points-impact}
# Extract points change from the summary or compute from result
if (!is.null(trade_result$summary)) {
  # Parse pts/week from summary structure
  pts_df <- data.frame(
    Category = c("Total", "Batting", "Pitching"),
    `Change (pts/wk)` = c(
      sprintf("%+.1f", trade_result$value_diff * 0.5),  # approximate
      "—", "—"
    ),
    stringsAsFactors = FALSE,
    check.names = FALSE
  )
  knitr::kable(pts_df, caption = "Projected Points Per Week Change")
}
```

---

## Salary Cap Impact

```{r salary-impact}
sal <- trade_result$salary_impact

salary_df <- data.frame(
  Metric = c("Net Salary Change", "New Team Total", "Salary Cap ($300)", "Status"),
  Value = c(
    sprintf("%+$d", sal$net_change),
    sprintf("$%.0f", sal$new_total),
    "$300",
    if (sal$cap_compliant) "✅ Compliant" else "❌ VIOLATION"
  ),
  stringsAsFactors = FALSE
)
knitr::kable(salary_df, caption = "Salary Cap Analysis")
```

---

## Keeper Implications

### Players You Would Give Up

```{r keeper-give}
if (length(trade_result$keeper_analysis$give) > 0) {
  give_keeper <- do.call(rbind, lapply(names(trade_result$keeper_analysis$give), function(pname) {
    info <- trade_result$keeper_analysis$give[[pname]]
    data.frame(
      Player = pname,
      `Current Salary` = sprintf("$%d", info$current_salary),
      `Year 1` = sprintf("$%d", info$escalation_path[1]),
      `Year 2` = sprintf("$%d", info$escalation_path[2]),
      `Year 3` = sprintf("$%d", info$escalation_path[3]),
      `3-Yr Keeper Value` = sprintf("$%.1f", info$keeper_value_3yr),
      `Keeper Eligible` = if (info$keeper_eligible) "Yes" else "No",
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }))
  knitr::kable(give_keeper, caption = "Keeper Analysis — Players Given")
} else {
  cat("*No keeper data available for players given.*")
}
```

### Players You Would Receive

```{r keeper-receive}
if (length(trade_result$keeper_analysis$receive) > 0) {
  receive_keeper <- do.call(rbind, lapply(names(trade_result$keeper_analysis$receive), function(pname) {
    info <- trade_result$keeper_analysis$receive[[pname]]
    data.frame(
      Player = pname,
      `Current Salary` = sprintf("$%d", info$current_salary),
      `Year 1` = sprintf("$%d", info$escalation_path[1]),
      `Year 2` = sprintf("$%d", info$escalation_path[2]),
      `Year 3` = sprintf("$%d", info$escalation_path[3]),
      `3-Yr Keeper Value` = sprintf("$%.1f", info$keeper_value_3yr),
      `Keeper Eligible` = if (info$keeper_eligible) "Yes" else "No",
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }))
  knitr::kable(receive_keeper, caption = "Keeper Analysis — Players Received")
} else {
  cat("*No keeper data available for players received.*")
}
```

---

## Positional Impact

```{r positional-impact}
pos_impact <- trade_result$positional_impact

if (length(pos_impact$improved) > 0 || length(pos_impact$weakened) > 0 ||
    length(pos_impact$overfilled) > 0 || length(pos_impact$unfilled) > 0) {
  
  pos_df <- data.frame(
    Category = c("Positions Improved", "Positions Weakened",
                 "Over-filled", "Left Unfilled"),
    Positions = c(
      if (length(pos_impact$improved) > 0) paste(pos_impact$improved, collapse = ", ") else "None",
      if (length(pos_impact$weakened) > 0) paste(pos_impact$weakened, collapse = ", ") else "None",
      if (length(pos_impact$overfilled) > 0) paste(pos_impact$overfilled, collapse = ", ") else "None",
      if (length(pos_impact$unfilled) > 0) paste(pos_impact$unfilled, collapse = ", ") else "None"
    ),
    stringsAsFactors = FALSE
  )
  knitr::kable(pos_df, caption = "Positional Impact Summary")
} else {
  cat("*No significant positional changes.*")
}
```

---

## Prospect Analysis

```{r prospect-analysis, results='asis'}
if (length(trade_result$prospect_analysis) > 0) {
  for (pname in names(trade_result$prospect_analysis)) {
    prospect <- trade_result$prospect_analysis[[pname]]
    cat(sprintf("### %s\n\n", pname))
    cat(sprintf("- **Prospect Tier:** %s\n", toupper(prospect$prospect_tier)))
    cat(sprintf("- **5-Year Keeper Value (NPV):** $%.1f\n", prospect$keeper_value_5yr))
    cat(sprintf("- **Current Salary Track:** $%d (Year %d)\n",
                prospect$salary_track$current_salary, prospect$salary_track$track_year))
    
    if (!is.null(prospect$draft_pedigree)) {
      dp <- prospect$draft_pedigree
      cat(sprintf("- **Draft Pedigree:** Round %d, Pick %d",
                  dp$round, dp$pick))
      if (!is.na(dp$signing_bonus)) {
        cat(sprintf(" ($%s bonus)", format(dp$signing_bonus, big.mark = ",")))
      }
      if (!is.na(dp$school)) {
        cat(sprintf(" — %s", dp$school))
      }
      cat("\n")
    } else {
      cat("- **Draft Pedigree:** Not found in draft data\n")
    }
    cat("\n")
  }
} else {
  cat("*No minor league players involved in this trade.*\n")
}
```

---

## Competitive Context

```{r competitive-context}
if (!is.null(standings)) {
  context <- assess_competitive_context(params$my_team, standings)
  
  context_df <- data.frame(
    Metric = c("Team Mode", "Playoff Probability", "Division Rank", "Games Back"),
    Value = c(
      toupper(context$mode),
      sprintf("%.0f%%", context$playoff_prob * 100),
      if (!is.na(context$division_rank)) as.character(context$division_rank) else "N/A",
      if (!is.na(context$games_back)) sprintf("%.1f", context$games_back) else "N/A"
    ),
    stringsAsFactors = FALSE
  )
  knitr::kable(context_df, caption = "Team Competitive Context")
  
  cat("\n")
  if (context$mode == "contend") {
    cat("*As a contender, prioritize immediate points impact over long-term keeper value.*\n")
  } else if (context$mode == "rebuild") {
    cat("*As a rebuilding team, prioritize keeper value and prospect accumulation.*\n")
  } else {
    cat("*In the middle — weigh both current-year impact and keeper value.*\n")
  }
}
```

---

*Report generated by the La-Z-Boyz Fantasy Baseball Decision Helper*
