---
title: "Player Valuation Report"
subtitle: "La-Z-Boyz of Summer Fantasy Baseball"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    df_print: kable
params:
  top_n: 50
  position_filter: NULL
date: "`r format(Sys.time(), '%B %d, %Y at %I:%M %p')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-dependencies}
# Source analysis modules
source(here::here("R/utils/scoring.R"))
source(here::here("R/utils/keeper_value.R"))
source(here::here("R/utils/salary_rules.R"))
source(here::here("R/utils/serialization.R"))
source(here::here("R/analysis/player_valuation.R"))
```

```{r load-data}
# Load cached data
valuations_path <- here::here("data/cache/valuations.rds")
valuations <- if (file.exists(valuations_path)) load_rds_safe(valuations_path) else NULL

rosters_path <- here::here("data/cache/rosters.rds")
rosters <- if (file.exists(rosters_path)) load_rds_safe(rosters_path) else NULL

projections_path <- here::here("data/cache/projections.rds")
projections <- if (file.exists(projections_path)) load_rds_safe(projections_path) else NULL
```

---

## Report Metadata

| Field | Value |
|-------|-------|
| **Analysis Type** | Player Valuation & Rankings |
| **Generated** | `r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')` |
| **Top N Players** | `r params$top_n` |
| **Position Filter** | `r if (is.null(params$position_filter)) "All" else params$position_filter` |
| **Salary Pool** | $4,160 (16 teams × $260) |

---

```{r data-check, results='asis'}
if (is.null(valuations)) {
  cat("## ❌ Valuations Not Available\n\n")
  cat("Run the full valuation pipeline to generate player values:\n\n")
  cat("```r\n")
  cat("source('R/analysis/player_valuation.R')\n")
  cat("valuations <- run_valuation_pipeline(projections)\n")
  cat("save_rds_with_metadata(valuations, 'data/cache/valuations.rds', 'valuation_pipeline')\n")
  cat("```\n")
  knitr::knit_exit()
}
```

## Overall Rankings

```{r overall-rankings}
# Apply position filter if specified
display_vals <- valuations

if (!is.null(params$position_filter)) {
  display_vals <- display_vals[
    toupper(display_vals$position) == toupper(params$position_filter), , drop = FALSE
  ]
}

# Sort by dollar value descending and take top N
display_vals <- display_vals[order(-display_vals$dollar_value), ]
top_players <- head(display_vals, params$top_n)

# Select display columns
display_cols <- intersect(
  c("player_name", "position", "proj_pts_per_week", "dollar_value",
    "salary", "surplus_value", "positional_scarcity"),
  names(top_players)
)
rankings_table <- top_players[, display_cols, drop = FALSE]

# Format column names
col_names <- c(
  player_name = "Player",
  position = "Pos",
  proj_pts_per_week = "Pts/Wk",
  dollar_value = "$ Value",
  salary = "Salary",
  surplus_value = "Surplus",
  positional_scarcity = "Scarcity"
)
names(rankings_table) <- col_names[names(rankings_table)]

# Round numeric columns
if ("Pts/Wk" %in% names(rankings_table)) {
  rankings_table$`Pts/Wk` <- round(rankings_table$`Pts/Wk`, 1)
}
if ("$ Value" %in% names(rankings_table)) {
  rankings_table$`$ Value` <- round(rankings_table$`$ Value`, 1)
}
if ("Surplus" %in% names(rankings_table)) {
  rankings_table$Surplus <- round(rankings_table$Surplus, 1)
}
if ("Scarcity" %in% names(rankings_table)) {
  rankings_table$Scarcity <- round(rankings_table$Scarcity, 2)
}

knitr::kable(rankings_table, caption = sprintf("Top %d Players by Dollar Value", params$top_n),
             row.names = FALSE)
```

---

## Positional Rankings

```{r positional-rankings, results='asis'}
positions <- c("C", "1B", "2B", "3B", "SS", "OF", "SP", "RP")

for (pos in positions) {
  pos_players <- valuations[toupper(valuations$position) == pos, , drop = FALSE]
  if (nrow(pos_players) == 0) next
  
  pos_players <- pos_players[order(-pos_players$dollar_value), ]
  pos_top <- head(pos_players, 10)
  
  pos_display <- pos_top[, intersect(
    c("player_name", "proj_pts_per_week", "dollar_value", "salary", "surplus_value"),
    names(pos_top)
  ), drop = FALSE]
  
  names(pos_display) <- c("Player", "Pts/Wk", "$ Value", "Salary", "Surplus")[
    seq_len(ncol(pos_display))
  ]
  
  # Round numerics
  for (col in names(pos_display)) {
    if (is.numeric(pos_display[[col]])) {
      pos_display[[col]] <- round(pos_display[[col]], 1)
    }
  }
  
  cat(sprintf("\n### %s — Top 10\n\n", pos))
  print(knitr::kable(pos_display, row.names = FALSE))
  cat("\n")
}
```

---

## Surplus Value Leaders

```{r surplus-leaders}
if ("surplus_value" %in% names(valuations)) {
  surplus_ranked <- valuations[!is.na(valuations$surplus_value), ]
  surplus_ranked <- surplus_ranked[order(-surplus_ranked$surplus_value), ]
  surplus_top <- head(surplus_ranked, 20)
  
  surplus_display <- surplus_top[, intersect(
    c("player_name", "position", "dollar_value", "salary", "surplus_value"),
    names(surplus_top)
  ), drop = FALSE]
  
  names(surplus_display) <- c("Player", "Pos", "$ Value", "Salary", "Surplus")[
    seq_len(ncol(surplus_display))
  ]
  
  for (col in names(surplus_display)) {
    if (is.numeric(surplus_display[[col]])) {
      surplus_display[[col]] <- round(surplus_display[[col]], 1)
    }
  }
  
  knitr::kable(surplus_display, caption = "Top 20 Surplus Value Players (Best Bargains)",
               row.names = FALSE)
}
```

---

## Keeper Value Rankings

```{r keeper-rankings}
if ("keeper_value_3yr" %in% names(valuations)) {
  keeper_ranked <- valuations[!is.na(valuations$keeper_value_3yr), ]
  keeper_ranked <- keeper_ranked[order(-keeper_ranked$keeper_value_3yr), ]
  keeper_top <- head(keeper_ranked, 20)
  
  keeper_display <- keeper_top[, intersect(
    c("player_name", "position", "salary", "dollar_value",
      "keeper_value_3yr", "keeper_value_5yr", "is_minor_contract"),
    names(keeper_top)
  ), drop = FALSE]
  
  col_names_keeper <- c(
    player_name = "Player",
    position = "Pos",
    salary = "Salary",
    dollar_value = "$ Value",
    keeper_value_3yr = "3-Yr Keeper",
    keeper_value_5yr = "5-Yr Keeper",
    is_minor_contract = "Minor?"
  )
  names(keeper_display) <- col_names_keeper[names(keeper_display)]
  
  for (col in names(keeper_display)) {
    if (is.numeric(keeper_display[[col]])) {
      keeper_display[[col]] <- round(keeper_display[[col]], 1)
    }
  }
  
  knitr::kable(keeper_display,
               caption = "Top 20 Keeper Value Players (3-Year NPV Surplus)",
               row.names = FALSE)
}
```

---

## Valuation Summary Statistics

```{r summary-stats}
summary_df <- data.frame(
  Metric = c(
    "Total Players Valued",
    "Players Above Replacement",
    "Mean Dollar Value",
    "Median Dollar Value",
    "Total Dollar Pool Distributed",
    "Mean Surplus Value (rostered)",
    "Players on Minor Contracts"
  ),
  Value = c(
    nrow(valuations),
    sum(valuations$dollar_value > 0, na.rm = TRUE),
    sprintf("$%.1f", mean(valuations$dollar_value, na.rm = TRUE)),
    sprintf("$%.1f", median(valuations$dollar_value, na.rm = TRUE)),
    sprintf("$%s", format(round(sum(valuations$dollar_value[valuations$dollar_value > 0],
                                     na.rm = TRUE)), big.mark = ",")),
    if ("surplus_value" %in% names(valuations)) {
      sprintf("$%.1f", mean(valuations$surplus_value, na.rm = TRUE))
    } else { "N/A" },
    if ("is_minor_contract" %in% names(valuations)) {
      sum(valuations$is_minor_contract == TRUE, na.rm = TRUE)
    } else { "N/A" }
  ),
  stringsAsFactors = FALSE
)
knitr::kable(summary_df, caption = "Valuation Engine Summary")
```

---

## Recommendations

```{r recommendations, results='asis'}
cat("### Key Takeaways\n\n")

# Identify most undervalued rostered players (highest surplus)
if ("surplus_value" %in% names(valuations) && !is.null(rosters)) {
  rostered_names <- tolower(rosters$player_name)
  rostered_vals <- valuations[tolower(valuations$player_name) %in% rostered_names, ]
  
  if (nrow(rostered_vals) > 0) {
    top_surplus <- rostered_vals[order(-rostered_vals$surplus_value), ]
    top3 <- head(top_surplus, 3)
    cat("**Best Value Players in the League:**\n\n")
    for (i in seq_len(nrow(top3))) {
      cat(sprintf("- %s (%s): $%.0f value on $%d salary = **$%.0f surplus**\n",
                  top3$player_name[i], top3$position[i],
                  top3$dollar_value[i], top3$salary[i], top3$surplus_value[i]))
    }
    cat("\n")
  }
}

# Identify most overvalued (negative surplus)
if ("surplus_value" %in% names(valuations) && !is.null(rosters)) {
  if (nrow(rostered_vals) > 0) {
    bottom_surplus <- rostered_vals[order(rostered_vals$surplus_value), ]
    bottom3 <- head(bottom_surplus, 3)
    cat("**Most Overpaid Players in the League:**\n\n")
    for (i in seq_len(nrow(bottom3))) {
      cat(sprintf("- %s (%s): $%.0f value on $%d salary = **$%.0f surplus**\n",
                  bottom3$player_name[i], bottom3$position[i],
                  bottom3$dollar_value[i], bottom3$salary[i], bottom3$surplus_value[i]))
    }
    cat("\n")
  }
}
```

---

*Report generated by the La-Z-Boyz Fantasy Baseball Decision Helper*
