---
title: "Weekly Decision Summary"
subtitle: "La-Z-Boyz of Summer Fantasy Baseball"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    df_print: kable
params:
  my_team: "My Team"
  week_number: 1
  remaining_faab: 250
  weeks_remaining: 23
date: "`r format(Sys.time(), '%B %d, %Y at %I:%M %p')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load-dependencies}
# Source analysis modules
source(here::here("R/utils/scoring.R"))
source(here::here("R/utils/player_linker.R"))
source(here::here("R/utils/keeper_value.R"))
source(here::here("R/utils/salary_rules.R"))
source(here::here("R/utils/serialization.R"))
source(here::here("R/analysis/projection_model.R"))
source(here::here("R/analysis/player_valuation.R"))
source(here::here("R/analysis/trade_analyzer.R"))
source(here::here("R/analysis/waiver_recommender.R"))
source(here::here("R/analysis/roster_optimizer.R"))
```

```{r load-data}
# Load cached data
rosters <- load_rds_safe(here::here("data/cache/rosters.rds"))
standings <- load_rds_safe(here::here("data/cache/standings.rds"))
players <- load_rds_safe(here::here("data/cache/players.rds"))

# Load valuations if available (generated by valuation pipeline)
valuations_path <- here::here("data/cache/valuations.rds")
valuations <- if (file.exists(valuations_path)) load_rds_safe(valuations_path) else NULL

# Load projections if available
projections_path <- here::here("data/cache/projections.rds")
projections <- if (file.exists(projections_path)) load_rds_safe(projections_path) else NULL
```

---

## Report Metadata

| Field | Value |
|-------|-------|
| **Analysis Type** | Weekly Decision Summary |
| **Team** | `r params$my_team` |
| **Week** | `r params$week_number` |
| **Generated** | `r format(Sys.time(), '%Y-%m-%d %H:%M:%S %Z')` |
| **FAAB Remaining** | $`r params$remaining_faab` |
| **Weeks Remaining** | `r params$weeks_remaining` |

---

## Recommended Lineup

```{r lineup}
if (!is.null(rosters) && !is.null(projections)) {
  lineup <- optimize_lineup(
    my_team = params$my_team,
    rosters = rosters,
    projections = projections,
    injuries = players
  )
  
  active <- lineup[lineup$status == "active", c("slot", "player_name", "proj_pts_per_week")]
  names(active) <- c("Slot", "Player", "Proj Pts/Wk")
  knitr::kable(active, caption = "Optimal Starting Lineup")
} else {
  cat("*Projections or roster data not available. Run the projection pipeline first.*")
}
```

```{r lineup-summary}
if (exists("lineup")) {
  total_pts <- sum(lineup$proj_pts_per_week[lineup$status == "active"], na.rm = TRUE)
  unfilled <- sum(lineup$status == "unfilled")
  cat(sprintf("\n**Projected Team Total:** %.1f pts/week\n", total_pts))
  if (unfilled > 0) {
    cat(sprintf("\n⚠️ **%d slot(s) unfilled** — check roster eligibility.\n", unfilled))
  }
}
```

---

## FAAB Targets

```{r faab-targets}
if (!is.null(valuations) && !is.null(rosters)) {
  faab_recs <- recommend_faab(
    my_team = params$my_team,
    valuations = valuations,
    rosters = rosters,
    remaining_faab = params$remaining_faab,
    weeks_remaining = params$weeks_remaining
  )
  
  if (nrow(faab_recs$targets) > 0) {
    targets_display <- faab_recs$targets[, intersect(
      c("player_name", "position", "dollar_value", "effective_surplus",
        "suggested_bid", "approaching_promotion"),
      names(faab_recs$targets)
    ), drop = FALSE]
    names(targets_display) <- gsub("_", " ", names(targets_display))
    names(targets_display) <- tools::toTitleCase(names(targets_display))
    knitr::kable(targets_display, caption = "Top FAAB Targets (by Surplus Value)")
  } else {
    cat(sprintf("*%s*", faab_recs$message))
  }
} else {
  cat("*Valuations not available. Run the valuation pipeline first.*")
}
```

### Drop Candidates

```{r drop-candidates}
if (exists("faab_recs") && nrow(faab_recs$drop_candidates) > 0) {
  drops <- head(faab_recs$drop_candidates, 5)
  drops_display <- drops[, intersect(
    c("player_name", "position", "salary", "surplus_value"),
    names(drops)
  ), drop = FALSE]
  names(drops_display) <- c("Player", "Position", "Salary", "Surplus Value")
  knitr::kable(drops_display, caption = "Top Drop Candidates (Lowest Surplus)")
}
```

---

## Trade Opportunities

```{r trade-opportunities}
if (!is.null(valuations) && !is.null(rosters) && !is.null(standings)) {
  context <- assess_competitive_context(params$my_team, standings)
  
  cat(sprintf("**Team Mode:** %s (Playoff Probability: %.0f%%)\n\n",
              toupper(context$mode), context$playoff_prob * 100))
  
  if (!is.na(context$division_rank)) {
    cat(sprintf("**Division Rank:** %d | **Games Back:** %.1f\n\n",
                context$division_rank, context$games_back))
  }
  
  if (context$mode == "contend") {
    cat("*Strategy: Target players with high immediate points output. ")
    cat("Consider trading prospects for proven contributors.*\n")
  } else if (context$mode == "rebuild") {
    cat("*Strategy: Target undervalued keepers and prospects. ")
    cat("Consider trading veterans for young talent with surplus value.*\n")
  } else {
    cat("*Strategy: Balanced approach — evaluate each deal on its merits. ")
    cat("Slight bias toward keeper value if playoff path is unclear.*\n")
  }
} else {
  cat("*Standings or valuations data not available.*")
}
```

---

## Keeper Value Watchlist

```{r keeper-watchlist}
if (!is.null(valuations) && "keeper_value_3yr" %in% names(valuations)) {
  # Find my team's players with best keeper value
  if (!is.null(rosters)) {
    my_roster <- rosters[tolower(rosters$team_name) == tolower(params$my_team), ]
    my_vals <- valuations[tolower(valuations$player_name) %in% tolower(my_roster$player_name), ]
    
    if (nrow(my_vals) > 0 && "keeper_value_3yr" %in% names(my_vals)) {
      my_vals <- my_vals[order(-my_vals$keeper_value_3yr), ]
      keeper_display <- head(my_vals, 10)
      keeper_display <- keeper_display[, intersect(
        c("player_name", "position", "salary", "dollar_value",
          "surplus_value", "keeper_value_3yr", "keeper_value_5yr"),
        names(keeper_display)
      ), drop = FALSE]
      names(keeper_display) <- gsub("_", " ", names(keeper_display))
      names(keeper_display) <- tools::toTitleCase(names(keeper_display))
      knitr::kable(keeper_display,
                   caption = "Top Keeper Value Players (3-Year NPV Surplus)")
    } else {
      cat("*No keeper value data available for your roster.*")
    }
  }
} else {
  cat("*Keeper values not computed. Run the full valuation pipeline first.*")
}
```

---

## Recommendations Summary

```{r recommendations}
cat("### Key Actions This Week\n\n")

actions <- character(0)

if (exists("lineup")) {
  unfilled <- sum(lineup$status == "unfilled")
  if (unfilled > 0) {
    actions <- c(actions, sprintf("- ⚠️ Address %d unfilled lineup slot(s)", unfilled))
  }
}

if (exists("faab_recs") && nrow(faab_recs$targets) > 0) {
  top_target <- faab_recs$targets$player_name[1]
  top_bid <- faab_recs$suggested_bids[1]
  actions <- c(actions, sprintf("- 💰 FAAB: Bid $%d on %s", top_bid, top_target))
}

if (exists("context")) {
  if (context$mode == "rebuild") {
    actions <- c(actions, "- 🔄 Explore selling veterans for keeper-value prospects")
  } else if (context$mode == "contend") {
    actions <- c(actions, "- 🏆 Focus on maximizing weekly points — trade deadline approaching")
  }
}

if (length(actions) == 0) {
  cat("*No urgent actions identified this week. Hold steady.*\n")
} else {
  cat(paste(actions, collapse = "\n"))
}
```

---

*Report generated by the La-Z-Boyz Fantasy Baseball Decision Helper*
